---
title: "St Croix leatherback juvenile survival rate estimation"
output: html_notebook
---


```{r}
rm(list=ls())
library(rstan)
library(loo)
library(tidyverse)
#set.seed(1)
```


Estimating average juvenile survival rates of leatherbacks from the numbers of neophytes and hatchlings. Because hatchlings are counted with errors but not neophytes, the time is reversed. Current data are stored so that they are separated into four groups; 9, 10, 11, and 12 years ago. The current dataset stops at 2005. 

Set up the dataset first.
```{r}
# the number of neophytes 9 yeras ago:
# 1994:2005
N9 <- c(45, 38, 58, 45, 90, 
        62, 49, 17, 61, 14, 
        25, 31)
# the number of neophytes 10 yeras ago:
# 1993:2004
N10 <- c(38, 58, 45, 90, 
         62, 49, 17, 61, 14, 
         25, 31, 26)
# the number of neophytes 11 yeras ago:
# 1992:2003
N11 <- c(58, 45, 90, 
         62, 49, 17, 61, 14, 
         25, 31, 26, 30)
# the number of neophytes 12 yeras ago:
# 1991:2002
N12 <- c(45, 90, 
         62, 49, 17, 61, 14, 
         25, 31, 26, 30, 13)

Y <- length(N12)

N <- c(N9, 26, 30, 13)

# Hatchlings end at 1993 - 12 years prior to the most recent (2005)
# 1982:1993
H <- c(8063, 10013, 
       7647, 4464, 4052, 7096, 5010, 
       2281, 6791, 3961, 3019, 1838)

hatch.yrs <- 1982:1993
```

Set up the stan input. First model is q-dot and phi-dot, where the proportion of neophytes that mature at age x (q(xs), where x = 9, 10, 11, or 12) are assumed equal.  Annual average survival rates are also assumed equal. 

```{r}
data_list = list(N9 = N9, N10 = N10, N11 = N11, N12 = N12,
                 Y = Y, H = H)

#n.cores <- options(mc.cores = parallel::detectCores())
mod_q_phi <- stan("models/Model_q_phi.stan", 
                 data=data_list, 
                 chains = 5, 
                 warmup = 5000,
                 iter = 10000,
                 cores = getOption("mc.cores", 4),
                 control = list(adapt_delta = 0.99))

```

Extract the log-likelihood and run the loo.
```{r}
log_lik_q_phi <- extract_log_lik(mod_q_phi, merge_chains = FALSE)
rel_n_eff_q_phi <- relative_eff(exp(log_lik_q_phi))
loo(log_lik_q_phi, r_eff = rel_n_eff_q_phi, cores = 4)

```

To see trace plots, use taceplot.

```{r}
#traceplot(mod, pars = c("p[1,1]", "p[1,2]", "p[1,3]", "p[1,4]"))
traceplot(mod_q_phi, pars = "q")
```

```{r}
traceplot(mod_q_phi, pars = c("phi"))

```

To see a summary of estimated paramters use print.
```{r}
# to see Rhat and other stats, use print
print(mod_q_phi)

```

Plot the data and predicted to see how they compare.

```{r}
list_of_samples <- rstan::extract(mod_q_phi)
names(list_of_samples)

mus <- as.data.frame(list_of_samples$mu)
colnames(mus) <- 1993:1982
mus2 <- reshape2::melt(mus, value.name = "log_hatchlings") %>%
  transmute(year = variable, log_hatchlings = log_hatchlings)

mus2$year <- factor(mus2$year, c(1982:1993))

# mus <- as.data.frame(list_of_samples$mu)
# colnames(mus) <- 1982:1993
# mus2 <- reshape2::melt(mus, value.name = "log_hatchlings")

# then the observed counts here:
obs.H <- data.frame(year = hatch.yrs, H = rev(H))

p.q.phi <- ggplot() + geom_boxplot(data = mus2,
                        aes(x = year, 
                            y = log_hatchlings)) + 
  geom_point(data = obs.H,
             aes(x = as.factor(year), 
                 y = log(H)),
             color = "red", size = 3)+ 
  labs(x = "", y = "ln(hatchlings)")

ggsave(filename = "figures/logHatch_q_phi.png",
       device = "png", dpi = 600,
       plot = p.q.phi)

```

Make the surival rate to be time varying. 

```{r}
data_list = list(N9 = N9, N10 = N10, N11 = N11, N12 = N12,
                 Y = Y, H = H)

#n.cores <- options(mc.cores = parallel::detectCores())
mod_q_phit <- stan("models/Model_q_phit.stan", 
                 data=data_list, 
                 chains = 5, 
                 warmup = 5000,
                 iter = 10000,
                 cores = getOption("mc.cores", 4),
                 control = list(adapt_delta = 0.99))

```

Extract the log-likelihood and run the loo.
```{r}
log_lik_q_phit <- extract_log_lik(mod_q_phit, 
                                  merge_chains = FALSE)
rel_n_eff_q_phit <- relative_eff(exp(log_lik_q_phit))
loo(log_lik_q_phit, r_eff = rel_n_eff_q_phit, cores = 4)

```


To see trace plots, use taceplot.

```{r}
traceplot(mod_q_phit, pars = "q")
```

```{r}
traceplot(mod_q_phit, pars = c("phi"))

```

```{r}
traceplot(mod_q_phit, pars = "sigma")
```


To see a summary of estimated paramters use print.
```{r}
# to see Rhat and other stats, use print
print(mod_q_phit)

```

Plot the data and predicted to see how they compare.

```{r}
list_of_samples <- rstan::extract(mod_q_phit)
names(list_of_samples)

mus <- as.data.frame(list_of_samples$mu)
colnames(mus) <- 1993:1982
mus2 <- reshape2::melt(mus, value.name = "log_hatchlings") %>%
  transmute(year = variable, log_hatchlings = log_hatchlings)

mus2$year <- factor(mus2$year, c(1982:1993))

# mus <- as.data.frame(list_of_samples$mu)
# colnames(mus) <- 1982:1993
# mus2 <- reshape2::melt(mus, value.name = "log_hatchlings")

# then the observed counts here:
obs.H <- data.frame(year = hatch.yrs, H = rev(H))

p.q.phit <- ggplot() + geom_boxplot(data = mus2,
                                    aes(x = year, 
                                        y = log_hatchlings)) + 
  geom_point(data = obs.H,
             aes(x = as.factor(year), 
                 y = log(H)),
             color = "red", size = 3) + 
  labs(x = "", y = "ln(hatchlings)")

ggsave(filename = "figures/logHatch_q_phit.png",
       device = "png", dpi = 600,
       plot = p.q.phit)
```

Varying the survival rates add more spread in the sigma parameter... 

Try with time-variable q

```{r}
data_list = list(N = N,
                 Y = Y, H = H)

#n.cores <- options(mc.cores = parallel::detectCores())
mod_qt_phi <- stan("models/Model_qt_phi.stan", 
                   data=data_list, 
                   chains = 5, 
                   warmup = 5000,
                   iter = 10000,
                   cores = getOption("mc.cores", 4),
                   control = list(adapt_delta = 0.99))

```

Get the log likelihood and run the loo.

```{r}
log_lik_qt_phi <- extract_log_lik(mod_qt_phi, 
                                  merge_chains = FALSE)
rel_n_eff_qt_phi <- relative_eff(exp(log_lik_qt_phi))
loo(log_lik_qt_phi, r_eff = rel_n_eff_qt_phi, cores = 4)

```


To see trace plots, use taceplot.

```{r}
traceplot(mod_qt_phi, pars = c("q[1,1]", "q[1,2]", "q[1,3]", "q[1,4]"))

```

```{r}
traceplot(mod_qt_phi, pars = c("phi"))

```

To see a summary of estimated paramters use print.
```{r}
# to see Rhat and other stats, use print
print(mod_qt_phi)

```

Plot the data and predicted to see how they compare.

```{r}
list_of_samples <- rstan::extract(mod_qt_phi)
names(list_of_samples)

mus <- as.data.frame(list_of_samples$mu)
colnames(mus) <- 1993:1982
mus2 <- reshape2::melt(mus, value.name = "log_hatchlings") %>%
  transmute(year = variable, log_hatchlings = log_hatchlings)

mus2$year <- factor(mus2$year, c(1982:1993))

# mus <- as.data.frame(list_of_samples$mu)
# colnames(mus) <- 1982:1993
# mus2 <- reshape2::melt(mus, value.name = "log_hatchlings")

# then the observed counts here:
obs.H <- data.frame(year = hatch.yrs, H = rev(H))

p.qt.phi <- ggplot() + geom_boxplot(data = mus2,
                        aes(x = year, 
                            y = log_hatchlings)) + 
  geom_point(data = obs.H,
             aes(x = as.factor(year), 
                 y = log(H)),
             color = "red", size = 3) + 
  labs(x = "", y = "ln(hatchlings)")

ggsave(filename = "figures/logHatch_qt_phi.png",
       device = "png", dpi = 600,
       plot = p.qt.phi)

```

Next qt and phit

```{r}
data_list = list(N = N,
                 Y = Y, H = H)

#n.cores <- options(mc.cores = parallel::detectCores())
mod_qt_phit <- stan("models/Model_qt_phit.stan", 
                   data=data_list, 
                   chains = 5, 
                   warmup = 5000,
                   iter = 10000,
                   cores = getOption("mc.cores", 4),
                   control = list(adapt_delta = 0.99))

```

Get the log likelihood and run the loo.

```{r}
log_lik_qt_phit <- extract_log_lik(mod_qt_phit, 
                                  merge_chains = FALSE)
rel_n_eff_qt_phit <- relative_eff(exp(log_lik_qt_phit))
loo(log_lik_qt_phit, r_eff = rel_n_eff_qt_phit, cores = 4)

```


To see trace plots, use taceplot.

```{r}
traceplot(mod_qt_phit, pars = c("q[1,1]", "q[1,2]", "q[1,3]", "q[1,4]"))

```

```{r}
traceplot(mod_qt_phit, pars = c("sigma"))

```


```{r}
traceplot(mod_qt_phit, pars = c("phi"))

```

To see a summary of estimated paramters use print.
```{r}
# to see Rhat and other stats, use print
print(mod_qt_phit)

```

Plot the data and predicted to see how they compare.

```{r}
list_of_samples <- rstan::extract(mod_qt_phit)
names(list_of_samples)

mus <- as.data.frame(list_of_samples$mu)
colnames(mus) <- 1993:1982
mus2 <- reshape2::melt(mus, value.name = "log_hatchlings") %>%
  transmute(year = variable, log_hatchlings = log_hatchlings)

mus2$year <- factor(mus2$year, c(1982:1993))

# mus <- as.data.frame(list_of_samples$mu)
# colnames(mus) <- 1982:1993
# mus2 <- reshape2::melt(mus, value.name = "log_hatchlings")

# then the observed counts here:
obs.H <- data.frame(year = hatch.yrs, H = rev(H))

p.qt.phit <- ggplot() + geom_boxplot(data = mus2,
                        aes(x = year, 
                            y = log_hatchlings)) + 
  geom_point(data = obs.H,
             aes(x = as.factor(year), 
                 y = log(H)),
             color = "red", size = 3)+ 
  labs(x = "", y = "ln(hatchlings)")

ggsave(filename = "figures/logHatch_qt_phit.png",
       device = "png", dpi = 600,
       plot = p.qt.phit)

```

This model seems to have a too large of sigma estimates... 


