---
title: "Expanding the range of age at first reproduction in the q-phi model"
output: html_notebook
---


```{r message=FALSE, echo=FALSE}
rm(list=ls())
library(rstan)
library(loo)
library(tidyverse)

save.fig <- T
save.results <- T
#set.seed(10)
n.cores <- parallel::detectCores()
MCMC.setup <- list(warmup = 5000,
                   iter = 10000,
                   chains = ifelse(n.cores < 5, n.cores, 5))
```

Estimating average juvenile survival rates of leatherbacks from the numbers of neophytes and hatchlings. Because hatchlings are counted with errors but not neophytes, the time is reversed.

In the previous analysis, four groups were used as possible age at first reproduction:  9, 10, 11, and 12 years. The next chunk can adjust the age ranges. 

Set up the dataset first.
```{r}

dat.0 <- read.csv(file = "data/hatchling and neophytes.csv")

# maximum age of first reproduction
max.age <- 13  #12
# the number of age classes for the first reproduction 
n.ages <- 5   # the number of age groups, given the max above. 

min.age <- max.age - n.ages + 1

# set up the data
max.yr <- dat.0$Year[nrow(dat.0)] - max.age
H <- dat.0$Hatchlings[dat.0$Year <= max.yr]

min.yr <- dat.0$Year[1] + min.age
# the number of neophytes:
neophytes <- rev(dat.0$Neophytes[dat.0$Year >= min.yr])

Y <- length(H)  
hatch.yrs <- dat.0$Year[1:Y]

```

Set up the stan input. The proportion of neophytes that mature at age x (q(xs), where x = min.age, ..., max.age) are assumed time-dependent.  Annual average survival rates are assumed time-independent (phi(.)). 

```{r message=FALSE, echo=FALSE}
data_list = list(N = neophytes,
                 Y = Y, 
                 H = rev(H),
                 n_ages = n.ages,
                 max_age = max.age)

mod_qt_phi <- stan("models/Model_qt_phi_1.stan", 
                   data=data_list, 
                   chains = MCMC.setup$chains, 
                   warmup = MCMC.setup$warmup,
                   iter = MCMC.setup$iter,
                   cores = getOption("mc.cores", MCMC.setup$chains),
                   control = list(adapt_delta = 0.99))

```

Extract the log-likelihood and run the loo.
```{r}
log_lik_qt_phi <- extract_log_lik(mod_qt_phi, 
                                  merge_chains = FALSE)
rel_n_eff_qt_phi <- relative_eff(exp(log_lik_qt_phi))
loo(log_lik_qt_phi, r_eff = rel_n_eff_qt_phi, cores = 4)

```

To see trace plots, use taceplot.

```{r}
traceplot(mod_qt_phi, pars = c("q[1,1]", "q[1,2]", "q[1,3]")) #, "q[1,4]"))

```

```{r}
traceplot(mod_qt_phi, pars = c("phi"))

```

To see a summary of estimated paramters use print.
```{r}
# to see Rhat and other stats, use print
print(mod_qt_phi)

```

Plot the data and predicted to see how they compare.

```{r}
list_of_samples <- rstan::extract(mod_qt_phi)
names(list_of_samples)

mus <- as.data.frame(list_of_samples$mu)
colnames(mus) <- rev(hatch.yrs) # 1993:1982
mus2 <- reshape2::melt(mus, value.name = "log_hatchlings") %>%
  transmute(year = variable, log_hatchlings = log_hatchlings)

mus2$year <- factor(mus2$year, hatch.yrs)

# mus <- as.data.frame(list_of_samples$mu)
# colnames(mus) <- 1982:1993
# mus2 <- reshape2::melt(mus, value.name = "log_hatchlings")

# then the observed counts here:
obs.H <- data.frame(year = hatch.yrs, H = H)

p.qt.phi <- ggplot() + geom_boxplot(data = mus2,
                                    aes(x = year, 
                                        y = log_hatchlings)) + 
  geom_point(data = obs.H,
             aes(x = as.factor(year), 
                 y = log(H)),
             color = "red", size = 3)+ 
  labs(x = "", y = "ln(hatchlings)")

print(p.qt.phi)
if (save.fig)
  ggsave(filename = paste0("figures/logHatch_qt_phi_maxAge_", 
                           max.age, "_nAges_", n.ages, ".png"),
         device = "png", dpi = 600,
         plot = p.qt.phi)

out.list <- list(output = mod_qt_phi,
                 MCMC.setup = MCMC.setup,
                 data = data_list,
                 plot = p.qt.phi)

if (save.results)
  write_rds(out.list, path = paste0("RData/DcJuvePhi_qt_phi_maxAge_", 
                                    max.age, "_nAges_", n.ages, ".rds"))


```


The plot suggests it's working but the loo results are not very satisfying... 
