---
title: "Expanding the range of age at first reproduction in the q-phi model"
output: html_notebook
---


```{r}
rm(list=ls())
library(rstan)
library(loo)
library(tidyverse)
#set.seed(1)
```

Estimating average juvenile survival rates of leatherbacks from the numbers of neophytes and hatchlings. Because hatchlings are counted with errors but not neophytes, the time is reversed.

In the previous analysis, four groups were used as possible age at first reproduction:  9, 10, 11, and 12 years. This analysis extends it to younger ages so that the average AFR may be estimated. 

Set up the dataset first.
```{r}

dat.0 <- read.csv(file = "data/hatchling and neophytes.csv")

max.yr <- dat.0$Year[nrow(dat.0)] - 12
hatchlings <- dat.0$Hatchlings[dat.0$Year <= max.yr]

min.yr <- dat.0$Year[1] + 6
# the number of neophytes:
neophytes <- rev(dat.0$Neophytes[dat.0$Year >= min.yr])

Y <- length(hatchlings)  
hatch.yrs <- dat.0$Year[1:Y]
```

Set up the stan input. First model is q-dot and phi-dot, where the proportion of neophytes that mature at age x (q(xs), where x = 9, 10, 11, or 12) are assumed equal.  Annual average survival rates are also assumed equal. 

```{r}
data_list = list(N = neophytes,
                 Y = Y, 
                 H = hatchlings,
                 n_ages = 6)

#n.cores <- options(mc.cores = parallel::detectCores())
mod_q_phi <- stan("models/Model_q_phi_1.stan", 
                 data=data_list, 
                 chains = 5, 
                 warmup = 5000,
                 iter = 10000,
                 cores = getOption("mc.cores", 4),
                 control = list(adapt_delta = 0.99))

```

Extract the log-likelihood and run the loo.
```{r}
log_lik_q_phi <- extract_log_lik(mod_q_phi, merge_chains = FALSE)
rel_n_eff_q_phi <- relative_eff(exp(log_lik_q_phi))
loo(log_lik_q_phi, r_eff = rel_n_eff_q_phi, cores = 4)

```

To see trace plots, use taceplot.

```{r}
#traceplot(mod, pars = c("p[1,1]", "p[1,2]", "p[1,3]", "p[1,4]"))
traceplot(mod_q_phi, pars = "q")
```

```{r}
traceplot(mod_q_phi, pars = c("phi"))

```

To see a summary of estimated paramters use print.
```{r}
# to see Rhat and other stats, use print
print(mod_q_phi)

```

Plot the data and predicted to see how they compare.

```{r}
list_of_samples <- rstan::extract(mod_q_phi)
names(list_of_samples)

mus <- as.data.frame(list_of_samples$mu)
colnames(mus) <- 1993:1982
mus2 <- reshape2::melt(mus, value.name = "log_hatchlings") %>%
  transmute(year = variable, log_hatchlings = log_hatchlings)

mus2$year <- factor(mus2$year, c(1982:1993))

# mus <- as.data.frame(list_of_samples$mu)
# colnames(mus) <- 1982:1993
# mus2 <- reshape2::melt(mus, value.name = "log_hatchlings")

# then the observed counts here:
obs.H <- data.frame(year = hatch.yrs, H = rev(H))

p.q.phi <- ggplot() + geom_boxplot(data = mus2,
                        aes(x = year, 
                            y = log_hatchlings)) + 
  geom_point(data = obs.H,
             aes(x = as.factor(year), 
                 y = log(H)),
             color = "red", size = 3)+ 
  labs(x = "", y = "ln(hatchlings)")

ggsave(filename = "figures/logHatch_q_phi.png",
       device = "png", dpi = 600,
       plot = p.q.phi)

```