---
title: "Model comparison"
output: html_notebook
---

In this document, I compare models that were run for estimating the juvenile survival rates and age at first reproduction for the St Croix leatherback turtle females. A various sets of age groups were analyzed using DcJuvePhi_stan.Rmd. Outputs were saved in .rds files. Because data were the same for all analyses (a various combinations of maximum age and the number of age groups), they should be comparable using looic values. 

```{r}
rm(list=ls())
library(rstan)
library(loo)
library(tidyverse)
library(bayesplot)
library(ggplot2)

```

First find all results files and bring them in one at a time and name them correctly. This chunk requires a bit to run, turn on cache. 

```{r cache=TRUE}
files <- list.files(path = "RData/", 
                    pattern = "DcJuvePhi_maxAge_")

looic.estimate <- looic.SE <- matrix(nrow = length(files), ncol = 4)

k <- 1
for (k in 1:length(files)){
  
  run.results <- readRDS(file = paste0("RData/", files[k]))
  loos <- run.results$loo
  looic.estimate[k,] <- c(loos$q_phi$estimates["looic", "Estimate"],
                          loos$q_phit$estimates["looic", "Estimate"],
                          loos$qt_phi$estimates["looic", "Estimate"],
                          loos$qt_phit$estimates["looic", "Estimate"])

  looic.SE[k,] <- c(loos$q_phi$estimates["looic", "SE"],
                    loos$q_phit$estimates["looic", "SE"],
                    loos$qt_phi$estimates["looic", "SE"],
                    loos$qt_phit$estimates["looic", "SE"])
  
}
```

Minimum looic is at the last one with qt_phi (time-dependent proportions that mature at different ages and a constant survival rate). Looic values decrease steadily as the maxAge increases. The last one (as of January 2019) is 17 years with 5 age groups (13-17 years old). It's possible that with a longer dataset, we may get to go higher. The difference in looic values between q_phi and qt_phi models shrink with the maxAge. For example, at maxAge = 10, the difference was more than 11. The difference became less than 4 when maxAge = 17. The models with time-dependent survival rates were not as good according to looic values. 

```{r}
# Find the row and column indices of the minimum
best.model.idx <- which(looic.estimate == min(looic.estimate),
                        arr.ind = TRUE)

# we are really interested in the row index to bring back the output
best.results <- readRDS(file = paste0("RData/", files[best.model.idx[1]]))

# model name:
best.name <- files[best.model.idx[1]]

# and take a look at pareto k values
loos <- best.results$loo
looic <- c(loos$q_phi$estimates["looic", "Estimate"],
           loos$q_phit$estimates["looic", "Estimate"],
           loos$qt_phi$estimates["looic", "Estimate"],
           loos$qt_phit$estimates["looic", "Estimate"])
plot(loos[[which(looic == min(looic))]])
```

Some of them are okay and a few are not... 

Because of the higher maxAge value, the posterior of the annual survival rate is pretty high. 

```{r}
models <- best.results$mods
model.best <- models[looic == min(looic)][[1]]

draws.df <- as.data.frame(model.best)
bayesplot::mcmc_dens(draws.df, pars = "phi")
```

The proportions that mature at different ages for cohorts:

```{r}
bayesplot::mcmc_dens(draws.df, pars = c("q[1,1]", "q[1,2]", 
                                        "q[1,3]", "q[1,4]", "q[1,5]"))
```

```{r}
bayesplot::mcmc_dens(draws.df, pars = c("q[2,1]", "q[2,2]", 
                                        "q[2,3]", "q[2,4]", "q[2,5]"))
```

```{r}
bayesplot::mcmc_dens(draws.df, pars = c("q[3,1]", "q[3,2]", 
                                        "q[3,3]", "q[3,4]", "q[3,5]"))
```

None of these seem to be any different from the equal proportions... so, this probably does not do any better than a constant q values. Let's take a look at q_phi results. 

```{r}
model.second <- models[1][[1]]
# and take a look at pareto k values
plot(loos[[1]])

```

Not that much better than qt_phi... 

```{r}
draws.df <- as.data.frame(model.second)
bayesplot::mcmc_dens(draws.df, pars = "phi")
```

This posterior is a tiny bit wider but not that much. 

```{r}
bayesplot::mcmc_dens(draws.df, pars = c("q[1]", "q[2]", 
                                        "q[3]", "q[4]", "q[5]"))
```

And these look very similar to those from the qt_phi models. So, why not make things simpler? 

```{r}
q_phi_summary <- summary(model.second, 
                         pars = c("phi", "q"),
                         probs = c(0.025, 0.5, 0.975))$summary

print(q_phi_summary)
```

*NOte - the following values change with different runs... pull out automatically (1/31/2019)
So, here it is. The average survival rate of juveniles is 0.75 with 95% CI of 0.72 - 0.78. And the proportions that mature at ages 13 - 17 are 0.13 (0.003 - 0.46), 0.27 (0.01 - 0.66), 0.16 (0.005 - 0.50), 0.20 (0.01 - 0.56), and 0.24 (0.01 - 0.62), respectively.

Save the results to a file so the posterior simulation can be done.

```{r}
# Create an output list; stan.out is the output from stan, model.str is the "structure" of 
# the model with max age and number of age classes for age at first reproduction. These two 
# bits of information are needed for simulations

out.list <- list(stan.out = best.results,
                 model.str = best.name)
saveRDS(out.list, file = "RData/model_output_for_post_sim.RDS")
```




