---
title: "Posterior simulation"
output: html_notebook
---

This notebook contains posterior simulation for the analysis of neophytes and hatchlings, which provided estimates of the average juvenile annual survival rate and the proportions of those that mature at various ages. All analyses were conducted in DcJuvePhi_stan.Rmd and a model comparison was conducted in DcJuvePhi_stan_model_comparison.Rmd. In this analysis, I extract the best results from the model comparison and simulate the number of neophytes based on the observed number of hatchlings and posterior samples of survival rate and proportions of age at first reproduction. 

```{r}
rm(list=ls())
library(rstan)
library(loo)
library(tidyverse)
library(bayesplot)
library(ggplot2)
```

Load the results from the appropriate model. Also load the data. 
```{r}

best.results <- readRDS(file = "RData/model_output_for_post_sim.RDS")
dat.0 <- read.csv(file = "data/hatchling and neophytes.csv")

# remove NAs from the hatchling data
dat.1 <- na.omit(dat.0)

# define the following parameters based on model.out$model.str
max.age <- 17
n.ages <- 5

min.age <- max.age - n.ages + 1
ages <- c(min.age:max.age)
```

Now we extract the posterior samples from the Stan output.

```{r}
loos <- best.results$stan.out$loo
looic <- c(loos$q_phi$estimates["looic", "Estimate"],
           loos$q_phit$estimates["looic", "Estimate"],
           loos$qt_phi$estimates["looic", "Estimate"],
           loos$qt_phit$estimates["looic", "Estimate"])

models <- best.results$stan.out$mods
model.second.best <- models[rank(looic) == 2][[1]]
model.second.best.name <- model.second.best@model_name
draws.df <- as.data.frame(model.second.best)
```

Start simulation using the hatchling data. 
```{r}
min.yr <- min(dat.1$Year)
max.yr <- max(dat.1$Year)
max.sim.yr <- max.yr + max.age
neophytes <- data.frame(Year = NA,
                        Neophytes = NA,
                        sample = NA)

nrows.tmp <- length((min.yr + min.age) : max.sim.yr)
r <- y <- i <- k <- 1 # nrow(draws.df)

# take first 1000 or take a random sample of 1000:
set.seed(123)
idx <- sort(sample.int(n = nrow(draws.df),
                       size = 1000, replace = FALSE))

for (i in 1:length(idx)){
  tmp <- data.frame(Year = (min.yr + min.age) : max.sim.yr,
                    Neophytes = 0,
                    sample = idx[i])
  # these mature at ages min.age to max.age
  qs <- draws.df[idx[i], c("q[1]", "q[2]", "q[3]", "q[4]", "q[5]")]
  phi <- (draws.df[idx[i], "phi"]) ^ c(min.age:max.age)
  for (k in 1:nrow(dat.1)){
    n.mature <- qs * dat.1[k, "Hatchlings"] * phi
    yrs <- dat.1[k, "Year"] + c(min.age:max.age)
    for ( y in 1:length(yrs)){
      tmp[tmp$Year == yrs[y], "Neophytes"] <- tmp[tmp$Year == yrs[y], "Neophytes"] + n.mature[1, y]
    }
  }

  neophytes[r:(r+nrows.tmp-1), ] <- tmp
  r <- r + nrows.tmp
}

```

Then plot the results.

```{r}
dat.2 <- filter(dat.0, Year >= min(tmp$Year))
ggplot() + 
  geom_boxplot(data = neophytes,
             aes(x = as.factor(Year), 
                 y = log(Neophytes)))  +
  geom_point(dat = dat.2,
             aes(x = as.factor(Year), y = log(Neophytes)),
             color = "red") + 
  labs(x = "Year", y = "log(Neophytes)")
```

Not sure if I want to call this "good" or "bad". Because of the large maximum age at first reproduction (17), only 11 data points can be compared. A qualitative comparison between the observed and simulated data looks "okay"... but not super exciting either. 


What if we use the more complicated qt_phi output? That was the best model.

```{r}
model.best <- models[rank(looic) == 1][[1]]
model.best.name <- model.best@model_name
draws.df.qt_phi <- as.data.frame(model.best)

neophytes.qt_phi <- data.frame(Year = NA,
                               Neophytes = NA,
                               sample = NA)

r <- y <- i <- k <- 1 # nrow(draws.df)

# take first 1000 or take a random sample of 1000:
set.seed(123)
idx <- sort(sample.int(n = nrow(draws.df.qt_phi),
                       size = 1000, 
                       replace = FALSE))
max.yr <- max(dat.0$Year)
nrows.tmp <- length((min.yr + min.age) : (min.yr + min.age + nrow(dat.0)-12+3))
qs <- draws.df.qt_phi %>% select(contains("q["))
for (i in 1:length(idx)){
  tmp <- data.frame(Year = (min.yr + min.age) : (min.yr + min.age + nrow(dat.0)-12+3),
                    Neophytes = 0,
                    sample = idx[i])
  # these mature at ages min.age to max.age
  
  phi <- (draws.df.qt_phi[idx[i], "phi"]) ^ c(min.age:max.age)
  q_i <- qs[idx[i],]
  for (k in 1:(nrow(dat.0)-12)){   # 13 (13 - 1 for counting) is the minimum age at maturity: 13-17 yo. may change. 
    # Each cohort maturing at different ages
    n.mature <- unlist(c(q_i[paste0("q[", k, ",1]")] * dat.0[k, "Hatchlings"] * phi[5], 
                         q_i[paste0("q[", k, ",2]")] * dat.0[k, "Hatchlings"] * phi[4],
                         q_i[paste0("q[", k, ",3]")] * dat.0[k, "Hatchlings"] * phi[3],
                         q_i[paste0("q[", k, ",4]")] * dat.0[k, "Hatchlings"] * phi[2],
                         q_i[paste0("q[", k, ",5]")] * dat.0[k, "Hatchlings"] * phi[1]))
    
    for (y in 1:length(n.mature)){
      tmp[k + y - 1, "Neophytes"] <- tmp[k + y - 1, "Neophytes"] + n.mature[y]
    }
    
  }
  neophytes.qt_phi[r:(r+nrows.tmp-1), ] <- tmp
  r <- r + nrows.tmp
}

```


```{r}
dat.2 <- filter(dat.0, Year >= min(tmp$Year))
ggplot() + 
  geom_boxplot(data = neophytes.qt_phi,
             aes(x = as.factor(Year), 
                 y = log(Neophytes)))  +
  geom_point(dat = dat.2,
             aes(x = as.factor(Year), y = log(Neophytes)),
             color = "red") + 
  labs(x = "Year", y = "log(Neophytes)")
```

