---
title: "Model testing with simulations"
output: html_notebook
---

```{r}
rm(list=ls())
library(rstan)
library(loo)
library(tidyverse)
```


This document describes simulations and analysis of simulated data to check if the models and Stan code are correctly specified. In the simulations, the number of hatchlings are assumed to be exact. Fixed proportions are used to derive the number of females maturing at ages 9, 10, 11, and 12. Also, fixed annual survival rates are used to simulate the fate of these hatchlings over time. Derived data on the number of annual neophytes are analyzed with models developed in Stan (used to be in JAGS) to determine if the correct estimates return. 

p(x) is the proportion of hatchlings that mature at age x, where x = 9,10,11, and 12, currently. p(9) + p(10) + p(11) + p(12) = 1.0. 

```{r}
# setup the simulation parameters

# Here are the raw data.
dat.0 <- read.csv(file = "data/hatchling and neophytes.csv")

# proportions mature at ages 9, 10, 11, and 12
p.9 <- 0.1
p.10 <- 0.3
p.11 <- 0.4
p.12 <- 0.2

# avearge annual survival rate 
phi <- 0.6

# maximum age of first reproduction
max.age <- 12
# the number of age classes for the first reproduction 
n.ages <- 4   # 4 means 9, 10, 11, and 12 years old are possible AFR ages.

min.age <- max.age - n.ages + 1
```

Using four cohorts at a time, they mature into one neophyte group. This simulation assumes that the proporiton that mature at certain age does not change over time. For example, p.9, which is the proportion of hatchlings in a cohort that mature after 9 years is the same for all cohorts. 

```{r}
set.seed(10)
# set up the data
# To create T years of neophytes data, we need to have T+n.ages-1 years of hatchling data
# This is the opposite of simulation_test2.Rmd, in which hatchling data were
# simulated from the neophytes. 

# because we are making all neophytes numbers, we can use all hatchlings data:
hatchlings <- na.omit(dat.0$Hatchlings)
sim.length <- length(hatchlings) - n.ages + 1
neophytes <- vector(mode = "numeric", length = sim.length)

for (k in 1:sim.length) {
  neophytes[k] <- (hatchlings[k] * p.12) * (phi^12) + 
    (hatchlings[k+1] * p.10) * (phi^11) + 
    (hatchlings[k+2] * p.11) * (phi^10) + 
    (hatchlings[k+3] * p.12) * (phi^9) + rnorm(n=1, mean = 0, sd = 10)
}

# the model uses neophytes to estimate hatchlings, going backwards in time. So,
# we need to extract hatchlings that match with the neophytes. Also, neophytes 
# vector needs to be reversed in time. The way this was simulated, we have the 
# same length for neophytes and hatchlings but in reality, this can't work because
# we can't use predicted numbers as data. 

Y <- 12
H <- hatchlings[1:Y]
N <- rev(neophytes[1:(Y+n.ages-1)])
```


Once the data are created, fit the model to the data to see if we can get the "right" answer back.

```{r}
# N9 <- neophytes[4:(sim.length-3)]
# N10 <- neophytes[3:(sim.length-4)]
# N11 <- neophytes[2:(sim.length-5)]
# N12 <- neophytes[1:(sim.length-6)]
# 
# Y <- length(N12)
# # hatchling vector needs to be reversed because the simulation goes forward, but the model requires the backward data. 
# H <- rev(hatchlings[1:Y])
# data_list = list(N9 = N9, 
#                  N10 = N10, 
#                  N11 = N11, 
#                  N12 = N12,
#                  Y = Y, H = H)

data_list = list(N = N,
                 Y = Y, 
                 H = H,
                 n_ages = 4,
                 max_age = 12)
```

In the model, q is the proportion of neophytes in year y that hatched x years ago, where x is 9, 10, 11, or 12. These proportions add up to 1.0 when they are constant over time.   

Then run the stan models.  First is the simplest: q-dot, phi-dot.
```{r}
mod_q_phi <- stan("models/Model_q_phi_1.stan", 
                  data=data_list, 
                  chains = 5, 
                  warmup = 5000,
                  iter = 10000,
                  cores = getOption("mc.cores", 5),
                  control = list(adapt_delta = 0.99))

```

Extract log likelihood and compute loo pareto stats.
```{r}

log_lik_q_phi <- extract_log_lik(mod_q_phi, 
                                 merge_chains = FALSE)
rel_n_eff_q_phi <- relative_eff(exp(log_lik_q_phi))
loo(log_lik_q_phi, r_eff = rel_n_eff_q_phi, cores = 8)

```


Look at the chains.
```{r}
traceplot(mod_q_phi, pars = "q")
```

```{r}
traceplot(mod_q_phi, pars = c("phi"))
```

```{r}
traceplot(mod_q_phi, pars = "sigma")

```



Look at the summary statistics.
```{r}
print(mod_q_phi)
```

Then look at the posteriors of log-hatchling numbers.

```{r}
# Because the model takes the time backwards, the output time is reversed. 
# Need to flip the output back in the right order. 

list_of_samples <- rstan::extract(mod_q_phi)
names(list_of_samples)

mus <- as.data.frame(list_of_samples$mu)
colnames(mus) <- 1993:1982
mus2 <- reshape2::melt(mus, value.name = "log_hatchlings") %>%
  transmute(year = variable, log_hatchlings = log_hatchlings)

mus2$year <- factor(mus2$year, c(1982:1993))
  #arrange(desc(year))

# then the observed counts here:
hatch.yrs <- 1982:1993
obs.H <- data.frame(year = hatch.yrs, H = H)

ggplot() + geom_boxplot(data = mus2,
                        aes(x = year, 
                            y = log_hatchlings)) + 
  geom_point(data = obs.H,
             aes(x = as.factor(year), 
                 y = log(H)),
             color = "red", 
             size = 3)

```

Looks a bit better now. It'd be good to have tighter estimates but when data are not precise, it's not possible...?

Next model has time-specific q but constant phi. For this model, one N vector needs to be constructed. 

```{r}
# N <- neophytes[1:(sim.length-3)]
# Y <- length(N) - 3
# data_list = list(N = N, 
#                  Y = Y, H = H)

#n.cores <- options(mc.cores = parallel::detectCores())
mod_qt_phi <- stan("models/Model_qt_phi_1.stan", 
                   data=data_list, 
                   chains = 5, 
                   warmup = 5000,
                   iter = 10000,
                   cores = getOption("mc.cores", 4),
                   control = list(adapt_delta = 0.99))


```


Get the log likelihood and do the loo...

```{r}
log_lik_qt_phi <- extract_log_lik(mod_qt_phi, merge_chains = FALSE)
rel_n_eff_qt_phi <- relative_eff(exp(log_lik_qt_phi))
loo(log_lik_qt_phi, r_eff = rel_n_eff_qt_phi, cores = 4)

```

```{r}
traceplot(mod_qt_phi, pars = c("q[1,1]", "q[1,2]", "q[1,3]", "q[1,4]"))
```
```{r}

traceplot(mod_qt_phi, pars = c("phi"))
```

```{r}
traceplot(mod_qt_phi, pars = "sigma")

```

To see a summary of estimated paramters use print.
```{r}
# to see Rhat and other stats, use print
print(mod_qt_phi)

```


Plot the data and predicted to see how they compare.

```{r}
list_of_samples <- rstan::extract(mod_qt_phi)
names(list_of_samples)

mus <- as.data.frame(list_of_samples$mu)
colnames(mus) <- 1993:1982
mus2 <- reshape2::melt(mus, value.name = "log_hatchlings") %>%
  transmute(year = variable, log_hatchlings = log_hatchlings)

mus2$year <- factor(mus2$year, c(1982:1993))
 
# then the observed counts here:
hatch.yrs <- 1982:1993
obs.H <- data.frame(year = hatch.yrs, H = H)

ggplot() + geom_boxplot(data = mus2,
                        aes(x = year, 
                            y = log_hatchlings)) + 
  geom_point(data = obs.H,
             aes(x = as.factor(year), 
                 y = log(H)),
             color = "red", size = 3)


```

Looks like the model works fine for estimating the parameters. 
